---
title: "Final Project"
author: "We Wish We Knew How to Code"
date: "4/18/2019"
output: html_document
---
#Final Project
##Members: Ryan Li, Alex Fan, Grace Kang, Sunny Chung, Huiyan Zhang 

####During the busiest time of the year, where are the most popular destinations (both clicked and actually booked) for people living in/outside the US?
Most popular destination for:
- Foreigner travelers to US
- US to US
- US to Foreign Countries (Completed)

####Setup
```{r echo=FALSE, error=FALSE, message=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(stringr)
library(ggmap)
library(purrr)
library(ggthemes)
data <- read_delim("data.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
dest <- read_delim("dest.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
```
Siran, Please note that we've included readr, tidyverse, lubridate, stringr, ggmap, purrr for libarary, and loaded dataset data.txt and dest.txt above. The loading dataset took a long time so we just set $include=FALSE$ to make sure our computers didn't explode. 

####Manipulating Data
```{r}
#selecting necessary columns within data file
data1<-data %>%
  select(date_time, user_location_country, user_location_latitude, orig_destination_distance, cnt, 
         srch_destination_id, user_location_city, hotel_country, hotel_id, is_booking, user_location_longitude)
#selecting necessary columns within destination file
dest1 <- dest %>%
  select(srch_destination_id, srch_destination_name, srch_destination_latitude, srch_destination_longitude)

#joining two dataset using left join because there were multiple data for date_time column, so we used srch_destination_id
joined <- left_join(data1, dest1, key = "srch_destination_id")
```
####


####Identifying Busiest Time of the Year
##select the busiest time data
```{r}
##Ggplot used to find out the busiest season

#we are plotting in a ggplot where x axis represents time of the year, and y axis represents the number of bookings at that time of the year. From the plot, it is obvious to find out that although the count constantly fluctuates, summer(around July) is the most popular time of the year. Therefore, we are selectin the busiest time--summer for future observation. 
data1 %>% 
  filter(is_booking ==1)%>%
  ggplot(aes(date_time)) + 
  geom_freqpoly(binwidth = 86400)+ # 86400 seconds = 1 day
  xlab("Date")+
  ylab("Number of Bookings")+
  ggtitle("Number of Bookings")+
  theme_economist()

#Zoom in to have a better vision of the data we are observing. 
usbooking <- joined %>% 
  filter(date_time < ymd_hms("2015-08-30 00:00:00"))%>%
  filter(ymd_hms("2015-06-01 00:00:00") < date_time)%>%
  filter(is_booking == 1) %>%
  filter(user_location_country == "UNITED STATES OF AMERICA")%>%
  ggplot(aes(date_time))+ 
  geom_freqpoly(binwidth = 86400)+ # 86400 seconds = 1 day
  xlab("Date") +
  ylab("Number of Bookings") +
  ggtitle("Number of Bookings from the US") +
  theme_economist()
  
  
usbooking

```

First graph indicates that majority of the bookings were done during summer time, we approximately chose a time frame to look at datas from June 1st, 2015 to August 30th 2015.

####TOP 5 Destinations
```{r}
#Top 5 destinations booked by users in the US
popular_booked <- joined %>%
  filter(is_booking==1)%>%
  filter(user_location_country=="UNITED STATES OF AMERICA")%>%
  count(hotel_country)%>%
  arrange(desc(n))%>%
  top_n(5)
popular_booked %>%
  ggplot(aes(x=hotel_country, y=n)) + geom_col(fill="red")+
  labs(x="Country", y="Number", title="Top 5 Destinations for US Users") +
  theme_economist() + 
  scale_color_economist()+
  theme(axis.text.x=element_text(angle=30, size=8, vjust=0.8))

#Top 5 destinations booked by users outside the US 
popular_booked1 <- joined %>%
  filter(is_booking==1)%>%
  filter(!user_location_country=="UNITED STATES OF AMERICA")%>%
  count(hotel_country)%>%
  arrange(desc(n)) %>%
  top_n(5)
popular_booked1 %>%
  ggplot(aes(x=hotel_country, y=n)) + geom_col(fill="orange")+
  labs(x="Country", y="Number", title="Top 5 Destinations for Non-US Users")+
  theme_economist()+
  scale_color_economist()+
  theme(axis.text.x=element_text(angle=30, size=8, vjust=0.8))

#Top 5 US Destinations for US users (state)
us_us <- joined %>%
  filter(is_booking == 1, user_location_country=="UNITED STATES OF AMERICA", hotel_country=="UNITED STATES OF AMERICA")%>%
  separate(srch_destination_name, into=c("city", "state", "country"), sep=",")%>%
  filter(!is.na(state))%>%
  count(state)%>%
  arrange(desc(n)) %>%
  top_n(5)
  
  
us_us%>%
  ggplot(aes(x=state, y=n, color=state, fill=state)) + geom_col(fill="lightyellow")+
  labs(x="State", y="Number", title="Top 5 US Destinations for US Users")+
  theme_economist()+
  scale_color_economist()+
  theme(axis.text.x=element_text(angle=30, size=8, vjust=0.8)) 

#Top 5 US Destinations for non-US users (state)
us_nonus <- joined %>%
  filter(is_booking == 1, !user_location_country=="UNITED STATES OF AMERICA", hotel_country=="UNITED STATES OF AMERICA")%>%
  separate(srch_destination_name, into=c("city", "state", "country"), sep=",")%>%
  filter(!is.na(state))%>%
  count(state)%>%
  arrange(desc(n)) %>%
  top_n(5)
us_nonus%>%
  ggplot(aes(x=state, y=n, color=state, fill=state)) + geom_col(fill="lightpink")+
  labs(x="State", y="Number", title="Top 5 US Destinations for Non-US Users")+
  theme_economist()+
  scale_color_economist()+
  theme(axis.text.x=element_text(angle=30, size=8, vjust=0.8))   

```

The most popular destinations booked by users in the US are (in the order of most booked to least booked) USA, Canada, Mexico, Italy, and UK. The result for users outside the US is not so much different: USA, Canada, Germany, Italy, and UK. Looking at the `popular_clicked` data, which has information of users that didn't actually book hotels, we can see that 4 out of top 5 destinations searched/clicked on Expedia are big cities in the US. Note that United States have extremely dominant number of booking because people tend to go for domestic travel. 



####
```{r}
#bookings from united states
kill_me <- joined %>%
  filter(is_booking == 1, user_location_country %in% "UNITED STATES OF AMERICA") %>%
  filter(!is.na(user_location_latitude), !is.na(srch_destination_latitude))


# make a map of the world
kill_me$hotel_country[kill_me$hotel_country %in% "UNITED STATES OF AMERICA"] <- "USA"

map.world <- map_data("world") # create a layer of borders
map.world$region <- toupper(map.world$region)

#Removed USA from map because of very high number of bookings, making it an outlier
new_kill_me<-kill_me %>%
  filter(hotel_country!="USA") %>%
  group_by(hotel_country) %>%
  summarize(total_bookings=n())
summary(new_kill_me)

#Join map of world with new dataset with USA removed
country_boi<-left_join(map.world, new_kill_me, by=c("region"="hotel_country"))

summary(country_boi)

#Create map. Midpoint set based on default scale
country_boi %>%
ggplot() +
  geom_polygon(data = country_boi, aes(x = long, y = lat, group = group, 
                                     fill=total_bookings)) +
  scale_fill_gradient2(low = ("yellow"), mid = "red",
                       high = ("blue"), midpoint = 5000, space = "Lab",
                       na.value = "grey50", guide = "colourbar", aesthetics = "fill")+
  theme(axis.text = element_blank(), axis.title = element_blank())+
  labs(fill="Total Bookings")+
  ggtitle("Where in the World are People in the US Traveling to?")
```

This is the ggmap of popular booking area by the United States Expedia users. Since majority of the bookings were done within United States for United States users and we're interested in where is the most popular booking location outside US for US users, we excluded booking data within US. 


####Before the presentation
So far, we've only answered the question for most poplular location for US users to foreign countries. For the following week, we will be changing user location id to look at most popular places for US users and foreign users. If we have enough time, we will try to use shiny and plotly to change the time frame in ggmap to better show how it's different over different period of time.






```{r}
##Ggplot used to find out the busiest season

#we are plotting in a ggplot where x axis represents time of the year, and y axis represents the number of bookings at that time of the year. From the plot, it is obvious to find out that although the count constantly fluctuates, summer(around July) is the most popular time of the year. Therefore, we are selectin the busiest time--summer for future observation. 
data1 %>% 
  filter(is_booking ==1)%>%
  ggplot(aes(date_time)) + 
  geom_freqpoly(binwidth = 86400) # 86400 seconds = 1 day


#Zoom in to have a better vision of the data we are observing. 
ocbooking <- joined %>% 
  filter(date_time < ymd_hms("2015-08-30 00:00:00"))%>%
  filter(ymd_hms("2015-06-01 00:00:00") < date_time)%>%
  filter(is_booking == 1) %>%
  filter(user_location_country != "UNITED STATES OF AMERICA")%>%
  ggplot(aes(date_time))+ 
  geom_freqpoly(binwidth = 86400) + # 86400 seconds = 1 day 
  xlab("Date") +
  ylab("Number of Bookings") +
  ggtitle("Number of Bookings from the US")
  
ocbooking

```



```{r}
#Top 5 destinations booked by users outside of US
ocpopular_booked <- joined %>%
  filter(is_booking==1)%>%
  filter(user_location_country!="UNITED STATES OF AMERICA")%>%
  count(hotel_country)%>%
  arrange(desc(n))%>%
  top_n(5)
popular_booked %>%
  ggplot(aes(x=hotel_country, y=n)) + geom_col()

#Top 5 destinations booked by users outside the US 
popular_booked1 <- joined %>%
  filter(is_booking==1)%>%
  filter(!user_location_country=="UNITED STATES OF AMERICA")%>%
  count(hotel_country)%>%
  arrange(desc(n)) %>%
  top_n(5)
popular_booked1 %>%
  ggplot(aes(x=hotel_country, y=n)) + geom_col()

#Top 5 Desinations clicked, but not actually booked
popular_clicked <-joined %>%
  filter(is_booking==0)%>%
  count(srch_destination_name)%>%
  arrange(desc(n))%>%
  top_n(5)
popular_clicked%>%
  ggplot(aes(x=srch_destination_name, y=n)) + geom_col()

popular_booked #US users
popular_booked1 #outside-US users
popular_clicked #all 
```


```{r}
# make new dataset for US-US Mapping
topstatebooked <- joined %>%
  filter(is_booking==1, user_location_country=="UNITED STATES OF AMERICA", hotel_country =="UNITED STATES OF AMERICA" )%>%
  separate(srch_destination_name, into=c("city","state","country"), sep=", ") # acquire states of search destinations
topstatebooked <- topstatebooked %>%
  filter(country %in% "United States of America")

# Clean the state names
topstatebooked$state[topstatebooked$state == "IA"] <- "Iowa"
topstatebooked$state[topstatebooked$state == "ID"] <- "Idaho"
topstatebooked$state[topstatebooked$state == "IL"] <- "Illinois"
topstatebooked$state[topstatebooked$state == "IN"] <- "Indiana"
topstatebooked$state[topstatebooked$state == "KS"] <- "Kansas"
topstatebooked$state[topstatebooked$state == "KY"] <- "Kentucky"
topstatebooked$state[topstatebooked$state == "MA"] <- "Massachusetts"
topstatebooked$state[topstatebooked$state == "MD"] <- "Maryland"
topstatebooked$state[topstatebooked$state == "MN"] <- "Minnesota"
topstatebooked$state[topstatebooked$state == "MT"] <- "Montana"
topstatebooked$state[topstatebooked$state == "NE"] <- "Nebraska"
topstatebooked$state[topstatebooked$state == "NV"] <- "Nevada"
topstatebooked$state[topstatebooked$state == "OH"] <- "Ohio"
topstatebooked$state[topstatebooked$state == "Okalhoma"] <- "Oklahoma"
topstatebooked$state[topstatebooked$state == "OR"] <- "Oregon"
topstatebooked$state[topstatebooked$state == "SC"] <- "South Carolina"
topstatebooked$state[topstatebooked$state == "VA"] <- "Virginia"
topstatebooked$state[topstatebooked$state == "VT"] <- "Vermont"
topstatebooked$state[topstatebooked$state == "WA"] <- "Washington"
topstatebooked$state[topstatebooked$state == "WI"] <- "Wisconsin"
topstatebooked$state <- tolower(topstatebooked$state)
table(topstatebooked$state)

# Summary of total bookings by state
summary_states <- topstatebooked %>%
  group_by(state) %>%
  summarize(Total = n())

library(maps)
states <- map_data("state")

table(states$region)

# join to state map dataset and run 
states <- left_join(states, summary_states, by = c("region" = "state"))
ggplot(data = states) +
  geom_polygon(aes(x = long, y = lat, group = group, fill = Total), color ="black") +
  scale_fill_gradient(low = "#FFD5CC", high = "#FF2D00") +
  ggtitle("US-US Traveling") +
  theme(axis.text = element_blank(), axis.title = element_blank()) +
  labs(fill="Total Bookings")
  
```

Here is the shiny web app code as well.
```{r echo = TRUE, eval = FALSE}
#Shiny Application for QTM 151 Project
library(shiny)

states <- read.csv("/Users/alexfan/Desktop/Emory Compiled/Emory Sem 4/QTM 151/Lab Project/states.csv")
country_boi <- read.csv("/Users/alexfan/Desktop/Emory Compiled/Emory Sem 4/QTM 151/Lab Project/country.csv")
# Define UI for application that draws a histogram
ui <- fluidPage(
  titlePanel("Maps that are cool"),
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "maptype", label = "What Map would you like today?",
                  choices = c("US-Foreign", "US-US"), selected = "US-Foreign"
      )
    ),
  mainPanel(
    plotlyOutput("plot")
  )
)
)
# Define server logic required to draw a histogram
server <- function(input, output){
  output$plot <- renderPlotly({
    if(input$maptype == "US-US"){
      gg_plot_US <- ggplot(data = states) +
        geom_polygon(aes(x = long, y = lat, group = group, fill = Total), color ="black") +
        scale_fill_gradient(low = "#FFD5CC", high = "#FF2D00")
      
      gg1 <- ggplotly(gg_plot_US)
      gg1 <- layout(gg1, dragmode = "pan")
      gg1
      
    }else if(input$maptype == "US-Foreign"){
      gg_world_plot <- country_boi %>%
        ggplot() +
        geom_polygon(data = country_boi, aes(x = long, y = lat, group = group, 
                                             fill=total_bookings)) +
        scale_fill_gradient2(low = ("yellow"), mid = "red",
                             high = ("blue"), midpoint = 5000, space = "Lab",
                             na.value = "grey50", guide = "colourbar", aesthetics = "fill")+
        theme(axis.text = element_blank(), axis.title = element_blank())+
        labs(fill="Total Bookings")
      
      gg2 <- ggplotly(gg_world_plot)
      gg2 <- layout(gg2, dragmode = "pan" )
      gg2
    }}
  )
}

# Run the application 
shinyApp(ui = ui, server = server)
```























